---
title: "Communicating spatial frames with whole-body silent gesture"
subtitle: "Descriptive statistics"
author: "Analysis by AĆ"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmarkdown::html_document:
    theme: readable
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
    df_print: paged
    code_folding: hide
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is the second script in the spatial-frames pipeline (the first, `1_preliminary.Rmd`, covered the semantic-distance work on the referent set). It describes the referential-game data. It fits nothing and tests nothing.

**The study.** Dyads play a referential game in which one partner (the *producer*) communicates a target referent to the other (the *guesser*) through whole-body silent gesture, and the guesser selects from a 16-image response matrix. The design is fully within-dyad, 2 (size: element vs. complete) × 2 (distance: proximate vs. displaced); every dyad completes all four conditions, 16 trials per dyad (one production of each of the 16 fixed stimuli), with producer and guesser roles alternating between partners.

**Two dependent measures.** *Communicative success* (`correct_target`, 1/0) is whether the guesser picked the right referent. *Representation strategy* is coded per production, and here the data depart from the preregistration: rather than a single icon-or-index label, each production carries three independent 0/1 flags — `icon`, `index`, and an unregistered third strategy, `prop` — and they can co-occur (a production may be coded icon *and* index). The modelling script therefore tests H2a and H2b as two separate Bernoulli models (`icon ~ .` and `index ~ .`), a principled deviation forced by discovering co-occurrence after preregistration; the registered single-binary Model A is retained there as a robustness check on the subset coded as exactly one of the two.

**What this script does.** It reads the annotation sheet, cleans it, derives the design variables from `story_id`, builds the unique person keys the models need, reports counts / balance / distributions / co-occurrence, and exports a prepared data frame for the modelling script. Everything the interpretation prose cites is stored in a named object, so the text tracks the data.

**Working on simulated data.** Until the empirical annotation arrives, `data_file` points at the simulated sheet, so every number below is placeholder. Swap that one path for the real coder file and nothing else changes.

# Data preparation

## Source setup

```{r source setup, echo = TRUE, message=FALSE, warning = FALSE}

########## folders ##########
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

parentfolder <- dirname(getwd())

data          <- paste0(parentfolder, '/data/')
models        <- paste0(parentfolder, '/models/')
plots         <- paste0(parentfolder, '/plots/')
scripts       <- paste0(parentfolder, '/scripts/')

#################### packages ####################
# Core tidyverse for wrangling + plotting; janitor/skimr for structure checks.
library(tidyverse)
library(janitor)
library(skimr)
library(readxl)
library(knitr)

# Keep plots/readability consistent
theme_set(theme_minimal(base_size = 12))

# use all available cores (harmless here; matches project setup)
options(mc.cores = parallel::detectCores())

colorBlindBlack8  <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Figure and Table caption helpers (manual numbering for HTML):
outputFormat = opts_knit$get("rmarkdown.pandoc.to");
if( is.null(outputFormat) ) outputFormat = "";
capTabNo = 1; capFigNo = 1;
capTab = function(x){
  if(outputFormat == 'html'){
    x = paste0("**Table ",capTabNo,".** ",x,"")
    capTabNo <<- capTabNo + 1
  }; x
}
capFig = function(x){
  if(outputFormat == 'html'){
    x = paste0("**Figure ",capFigNo,".** ",x,"")
    capFigNo <<- capFigNo + 1
  }; x
}
```

## Load in data

```{r read data, echo=TRUE, message=FALSE, warning=FALSE}
# SWAP this one line for the real coder sheet when it arrives; if the real
# file has several coder sheets, add e.g. sheet = "coder_MS".
data_file <- paste0(data, "dataframe_consensus.xlsx")

df <- read_excel(data_file) %>%
  clean_names() %>%
  remove_empty(c("rows", "cols"))   # drops trailing empty rows / stray blank column

# --- FIX: pair11 participant mislabelling ----------------------------------- #
# The 11th dyad is p23 / p24; trials 1-2 were entered with stale IDs p21 / p22.
# Corrected here (not in the raw file) so the source sheet stays untouched and
# the correction is documented in the pipeline.
df <- df %>%
  mutate(
    participant = if_else(
      pair == "pair11",
      case_match(participant,
                 "p21" ~ "p23",
                 "p22" ~ "p24",
                 .default = participant),
      participant
    )
  )

# Peek at raw types and missingness
skim(df)
```

## Derive design variables and person keys

`story_id` encodes the design as `[size][distance][item]` (e/c = element/complete, p/d = proximate/displaced, 1-4 = item within cell), so `size` and `distance` are decoded straight from it. `story_id` is 1:1 with `trial_id` and `target`, so it also serves as the **item** for the by-item random effect (16 levels). The within-cell index (`item_num`, 1-4) is descriptive only and is *not* the modeling item.

`participant` is a slot *within* the dyad, so it is not a unique person id (slot 1 in one pair is a different person from slot 1 in another). We build a unique producer key by crossing `pair` × `participant`, and the guesser is the partner in the same dyad.

```{r derive, echo=TRUE, message=FALSE, warning=FALSE}
# --- decode the design from story_id (character at this point) ---
df <- df %>%
  mutate(
    size     = if_else(substr(story_id, 1, 1) == "e", "element",   "complete"),
    distance = if_else(substr(story_id, 2, 2) == "p", "proximate", "displaced"),
    item_num = substr(story_id, 3, 3)
  )

# --- integrity checks: exactly two participants per pair; 16 stimuli ---
n_per_pair <- df %>% distinct(pair, participant) %>% count(pair)
if (any(n_per_pair$n != 2))
  stop("Each pair must have exactly two participants - check the data.")
if (n_distinct(df$story_id) != 16)
  warning("Expected 16 unique stimuli in story_id.")

# --- guesser = the partner within the dyad (exactly two members per pair) ---
partner_lookup <- df %>%
  distinct(pair, participant) %>%
  group_by(pair) %>%
  mutate(guesser_slot = rev(participant)) %>%   # 2 rows per pair -> swaps them
  ungroup()

df <- df %>%
  left_join(partner_lookup, by = c("pair", "participant")) %>%
  mutate(
    producer = paste(pair, participant,  sep = "_"),   # unique producer id
    guesser  = paste(pair, guesser_slot, sep = "_")    # unique guesser id
  )
```

## Cleaning and types

```{r types, echo=TRUE, message=FALSE, warning=FALSE}
df <- df %>%
  mutate(
    pair        = factor(pair),
    participant = factor(participant),
    producer    = factor(producer),
    guesser     = factor(guesser),
    story_id    = factor(story_id),          # 16-level item for random effects
    target      = factor(target),
    size        = factor(size,     levels = c("element", "complete")),
    distance    = factor(distance, levels = c("proximate", "displaced")),
    item_num    = factor(item_num),          # 1-4 within-cell index (descriptive)
    trial       = as.integer(trial),
    trial_id    = as.integer(trial_id),
    correct_target = as.integer(correct_target),
    icon  = as.integer(icon),
    index = as.integer(index),
    prop  = as.integer(prop)
  )
```

## Exclusions

Following the preregistration, trials lost to recording failure (no recorded success) or uncodeable for strategy are excluded and their number reported; no outlier- or performance-based exclusions are applied.

```{r exclusions, echo=TRUE, message=FALSE, warning=FALSE}
df <- df %>%
  mutate(
    success_recorded = !is.na(correct_target),
    strategy_coded   = !(is.na(icon) & is.na(index) & is.na(prop))
  )

n_rows_raw   <- nrow(df)
n_no_success <- sum(!df$success_recorded)
n_uncodeable <- sum(!df$strategy_coded)

# Keep only fully codeable productions for all downstream summaries
df <- df %>% filter(success_recorded, strategy_coded)
n_rows_ok <- nrow(df)
```

Of `r n_rows_raw` recorded productions, `r n_no_success` lacked a recorded outcome and `r n_uncodeable` were uncodeable for strategy, leaving **`r n_rows_ok`** codeable productions for analysis.

# Descriptive statistics

## Sample and balance

These quantities are near-constant by design (the design is balanced and fully crossed), so they are reported as checks: they only move under exclusions.

```{r counts, echo=TRUE, message=FALSE, warning=FALSE}
n_pairs_obs    <- n_distinct(df$pair)
n_participants <- n_distinct(c(as.character(df$producer),
                              as.character(df$guesser)))

trials_per_pair   <- df %>% count(pair, name = "n_trials")
prod_per_producer <- df %>% count(producer, name = "n_productions")
prod_per_guesser  <- df %>% count(guesser,  name = "n_guesses")
prod_per_item     <- df %>% count(story_id, target, name = "n_productions")

min_item_n <- min(prod_per_item$n_productions)
max_item_n <- max(prod_per_item$n_productions)

trials_per_pair
```

```{r counts_items, echo=TRUE, message=FALSE, warning=FALSE}
prod_per_item %>% arrange(n_productions)
```

The data cover `r n_pairs_obs` dyads (`r n_participants` participants), with `r min(trials_per_pair$n_trials)`-`r max(trials_per_pair$n_trials)` trials per dyad and each producer contributing `r min(prod_per_producer$n_productions)`-`r max(prod_per_producer$n_productions)` productions. Productions per unique referent range from `r min_item_n` to `r max_item_n` (target: at least 20 per referent).

## Communicative success

```{r success, echo=TRUE, message=FALSE, warning=FALSE}
mean_success <- mean(df$correct_target)

success_by_size <- df %>%
  group_by(size) %>%
  summarise(n = n(), rate = mean(correct_target), .groups = "drop")

success_by_distance <- df %>%
  group_by(distance) %>%
  summarise(n = n(), rate = mean(correct_target), .groups = "drop")

success_by_cell <- df %>%
  group_by(size, distance) %>%
  summarise(n = n(),
            rate = mean(correct_target),
            se   = sqrt(rate * (1 - rate) / n), .groups = "drop")

# scalars for inline use
succ_element  <- success_by_size$rate[success_by_size$size == "element"]
succ_complete <- success_by_size$rate[success_by_size$size == "complete"]
succ_prox     <- success_by_distance$rate[success_by_distance$distance == "proximate"]
succ_disp     <- success_by_distance$rate[success_by_distance$distance == "displaced"]

success_by_cell
```

Overall, guessers identified the correct target on `r round(100 * mean_success, 1)`% of productions. By referent size, element targets were correct `r round(100 * succ_element, 1)`% of the time versus `r round(100 * succ_complete, 1)`% for complete targets (H1 predicts complete < element; formal testing is in the results script). By distance, proximate targets were correct `r round(100 * succ_prox, 1)`% versus `r round(100 * succ_disp, 1)`% for displaced (H2 predicts no difference on success).

```{r fig_success, echo=FALSE, fig.align='left', fig.cap=capFig("Communicative success rate by size and distance (bars are ±1 binomial SE)."), fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
ggplot(success_by_cell, aes(x = distance, y = rate, fill = size)) +
  geom_col(position = position_dodge(0.7), width = 0.6, alpha = 0.9) +
  geom_errorbar(aes(ymin = rate - se, ymax = rate + se),
                width = 0.15, position = position_dodge(0.7)) +
  scale_fill_manual(values = colorBlindBlack8[c(4, 7)]) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(x = "Distance", y = "Correct-target rate", fill = "Size") +
  theme_minimal()
```

## Participant variability

The models carry by-producer and by-guesser random effects. This is the raw spread they absorb.

```{r participant_var, echo=TRUE, message=FALSE, warning=FALSE}
producer_success <- df %>%
  group_by(id = producer) %>%
  summarise(role = "producer", n = n(),
            rate = mean(correct_target),
            se   = sqrt(rate * (1 - rate) / n), .groups = "drop")

guesser_success <- df %>%
  group_by(id = guesser) %>%
  summarise(role = "guesser", n = n(),
            rate = mean(correct_target),
            se   = sqrt(rate * (1 - rate) / n), .groups = "drop")

participant_success <- bind_rows(producer_success, guesser_success) %>%
  mutate(role = factor(role, levels = c("producer", "guesser")))

prod_rate_range  <- range(producer_success$rate)
guess_rate_range <- range(guesser_success$rate)
```

Producer success rates ranged from `r round(100 * prod_rate_range[1], 1)`% to `r round(100 * prod_rate_range[2], 1)`% and guesser rates from `r round(100 * guess_rate_range[1], 1)`% to `r round(100 * guess_rate_range[2], 1)`%. With a median of `r median(producer_success$n)` productions per participant these per-person estimates are noisy, which is what partial pooling addresses.

```{r fig_participant_var, echo=FALSE, fig.align='left', fig.cap=capFig("Per-participant communicative success rate, ranked within role (points ±1 binomial SE). Spread previews the by-producer and by-guesser random effects; wide bars reflect few productions per participant."), fig.height=5, fig.width=9, message=FALSE, warning=FALSE}
participant_success %>%
  group_by(role) %>%
  arrange(rate) %>%
  mutate(rank = row_number()) %>%
  ungroup() %>%
  ggplot(aes(x = rank, y = rate)) +
  geom_errorbar(aes(ymin = pmax(0, rate - se), ymax = pmin(1, rate + se)),
                width = 0, colour = "grey75") +
  geom_point(size = 1.8, colour = colorBlindBlack8[6]) +
  facet_wrap(~ role, scales = "free_x") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(x = "Participant (ranked by success rate)", y = "Correct-target rate") +
  theme_minimal()
```

## Success by referent

```{r item_success, echo=TRUE, message=FALSE, warning=FALSE}
success_by_item <- df %>%
  group_by(story_id, target, size, distance) %>%
  summarise(n = n(),
            rate = mean(correct_target),
            se   = sqrt(rate * (1 - rate) / n), .groups = "drop")

# scalars for inline use
hardest_item <- success_by_item %>% slice_min(rate, n = 1, with_ties = FALSE)
easiest_item <- success_by_item %>% slice_max(rate, n = 1, with_ties = FALSE)

success_by_item %>% arrange(rate)
```

The referent hardest to communicate was `r as.character(hardest_item$target)` (`r round(100 * hardest_item$rate, 1)`%), the easiest `r as.character(easiest_item$target)` (`r round(100 * easiest_item$rate, 1)`%). Spread of this size across the 16 fixed referents is what the by-item random effect absorbs.

```{r fig_item_success, echo=FALSE, fig.align='left', fig.cap=capFig("Communicative success rate per referent (points ±1 binomial SE). Colour = size (element / complete), shape = distance (dot = proximate, triangle = displaced). Each referent has a fixed size and distance, so the points fall into the four design cells; spread across the fixed 16-referent set motivates the by-item random effect."), fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
ggplot(success_by_item, aes(x = reorder(target, rate), y = rate)) +
  geom_errorbar(aes(ymin = pmax(0, rate - se), ymax = pmin(1, rate + se)),
                width = 0.25, colour = "grey70") +
  geom_point(aes(colour = size, shape = distance), size = 2.8) +
  scale_colour_manual(values = colorBlindBlack8[c(4, 7)]) +
  scale_shape_manual(values = c(proximate = 16, displaced = 17)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  coord_flip() +
  labs(x = "Referent (target)", y = "Correct-target rate",
       colour = "Size", shape = "Distance") +
  theme_minimal()
```

## Strategy use

```{r strategy, echo=TRUE, message=FALSE, warning=FALSE}
prop_icon  <- mean(df$icon)
prop_index <- mean(df$index)
prop_prop  <- mean(df$prop)

strategy_by_cell <- df %>%
  group_by(size, distance) %>%
  summarise(icon = mean(icon), index = mean(index), prop = mean(prop),
            n = n(), .groups = "drop")

# scalars: distance is the key axis for H2a/H2b
icon_prox  <- mean(df$icon[df$distance == "proximate"])
icon_disp  <- mean(df$icon[df$distance == "displaced"])
index_prox <- mean(df$index[df$distance == "proximate"])
index_disp <- mean(df$index[df$distance == "displaced"])

strategy_by_cell
```

Marginally, icon was used on `r round(100 * prop_icon, 1)`% of productions, index on `r round(100 * prop_index, 1)`%, and the unregistered prop on `r round(100 * prop_prop, 1)`%. Across the distance contrast that H2a/H2b concern, icon use went from `r round(100 * icon_prox, 1)`% (proximate) to `r round(100 * icon_disp, 1)`% (displaced), and index use from `r round(100 * index_prox, 1)`% (proximate) to `r round(100 * index_disp, 1)`% (displaced) - the pattern H2a (displaced → more icon) and H2b (proximate → more index) predict.

```{r fig_strategy, echo=FALSE, fig.align='left', fig.cap=capFig("Strategy use rate by strategy, faceted by size and distance."), fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
strategy_by_cell_long <- strategy_by_cell %>%
  pivot_longer(c(icon, index, prop), names_to = "strategy", values_to = "rate") %>%
  mutate(strategy = factor(strategy, levels = c("icon", "index", "prop")))

ggplot(strategy_by_cell_long, aes(x = strategy, y = rate, fill = strategy)) +
  geom_col(width = 0.7, alpha = 0.9) +
  facet_grid(size ~ distance) +
  scale_fill_manual(values = colorBlindBlack8[c(2, 3, 4)]) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Use rate", fill = "Strategy") +
  theme_minimal()
```

Article plot:

```{r}

# --- data: icon / index only ------------------------------------------------ #
strategy_by_cell_long <- strategy_by_cell %>%
  pivot_longer(c(icon, index), names_to = "strategy", values_to = "rate") %>%
  mutate(strategy = factor(strategy, levels = c("icon", "index")))

# --- publication figure ----------------------------------------------------- #
p_strategy <- ggplot(strategy_by_cell_long,
                     aes(x = strategy, y = rate, fill = strategy)) +
  geom_col(width = 0.65) +
  geom_text(aes(label = scales::percent(rate, accuracy = 1)),
            vjust = -0.6, size = 6.5, colour = "grey15") +
  facet_grid(
    size ~ distance,
    labeller = labeller(
      size     = c(element   = "Element",   complete  = "Complete"),
      distance = c(proximate = "Proximate", displaced = "Displaced")
    )
  ) +
  scale_fill_manual(values = colorBlindBlack8[c(2, 3)]) +
  scale_y_continuous(
    labels = scales::percent,
    breaks = seq(0, 1, 0.25),
    limits = c(0, 1.08),
    expand = expansion(mult = c(0, 0))
  ) +
  labs(x = NULL, y = "Use rate") +
  theme_minimal(base_size = 20) +
  theme(
    legend.position    = "none",
    axis.title.y       = element_text(margin = margin(r = 12)),
    axis.text          = element_text(colour = "black"),
    axis.text.x        = element_text(face = "bold"),
    strip.text         = element_text(face = "bold"),
    strip.text.y       = element_text(angle = -90),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.spacing      = grid::unit(1.4, "lines"),
    plot.background    = element_rect(fill = "white", colour = NA),
    plot.margin        = margin(12, 16, 12, 12)
  )

p_strategy

# --- export ----------------------------------------------------------------- #

ggsave(file.path(plots, "fig_strategy_icon_index.svg"),
       plot = p_strategy, device = svglite::svglite,
       width = 9, height = 7, units = "in", bg = "white")

ggsave(file.path(plots, "fig_strategy_icon_index.png"),
       plot = p_strategy, device = ragg::agg_png,
       width = 9, height = 7, units = "in", dpi = 600, bg = "white")
```


## Icon-index co-occurrence

The reason the strategy analysis splits into two Bernoullis: icon and index are not complementary. This is the model-free view of how often they co-occur, and whether co-occurrence tracks the design - the question the exploratory categorical model would formalise.

```{r cooccur, echo=TRUE, message=FALSE, warning=FALSE}
prop_both <- mean(df$icon == 1 & df$index == 1)

cooccur_tab <- df %>%
  count(icon, index, name = "n") %>%
  mutate(prop = n / sum(n))

# four-way configuration of the icon x index pattern (prop handled separately)
df <- df %>%
  mutate(strategy_combo = case_when(
    icon == 1 & index == 1 ~ "both",
    icon == 1 & index == 0 ~ "icon_only",
    icon == 0 & index == 1 ~ "index_only",
    icon == 0 & index == 0 ~ "neither"
  ) %>% factor(levels = c("neither", "icon_only", "index_only", "both")))

combo_by_cell <- df %>%
  count(size, distance, strategy_combo, name = "n") %>%
  group_by(size, distance) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

cooccur_tab
```

Icon and index were coded on the same production in `r round(100 * prop_both, 1)`% of cases. Whether that layering concentrates in particular cells is what the figure below shows and the categorical model would test. (Note: `neither` is defined on icon × index only, so it includes prop-only productions.)

```{r fig_cooccur, echo=FALSE, fig.align='left', fig.cap=capFig("Distribution of the icon x index configuration across the 2x2 design - the model-free view the exploratory categorical model formalises."), fig.height=5, fig.width=9, message=FALSE, warning=FALSE}
ggplot(combo_by_cell, aes(x = distance, y = prop, fill = strategy_combo)) +
  geom_col(width = 0.65, alpha = 0.9) +
  facet_wrap(~ size) +
  scale_fill_manual(values = colorBlindBlack8[c(1, 2, 3, 6)]) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Distance", y = "Share of productions", fill = "icon x index") +
  theme_minimal()
```

## Success by strategy configuration

```{r success_by_strategy, echo=TRUE, message=FALSE, warning=FALSE}
success_by_combo <- df %>%
  group_by(strategy_combo) %>%
  summarise(n = n(),
            rate = mean(correct_target),
            se   = sqrt(rate * (1 - rate) / n), .groups = "drop")

success_by_combo
```

This is descriptive of the `icon` and `index` covariates that enter Model B; it is not itself a test.

```{r fig_success_by_strategy, echo=FALSE, fig.align='left', fig.cap=capFig("Communicative success rate by icon x index configuration (points ±1 binomial SE). 'neither' includes prop-only productions."), fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
ggplot(success_by_combo, aes(x = strategy_combo, y = rate)) +
  geom_col(width = 0.6, fill = colorBlindBlack8[3], alpha = 0.9) +
  geom_errorbar(aes(ymin = pmax(0, rate - se), ymax = pmin(1, rate + se)),
                width = 0.15) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(x = "icon x index configuration", y = "Correct-target rate") +
  theme_minimal()
```

# Export prepared data

The prepared frame feeds `3_modeling.Rmd`. Saved as `.rds` to preserve factor levels (and any contrasts set later); a `.csv` copy is written for inspection. Redundant helper columns are dropped first.

```{r export, echo=TRUE, message=FALSE, warning=FALSE}
df <- df %>% select(-guesser_slot, -success_recorded, -strategy_coded)

saveRDS(df,  paste0(data, "df_prepared.rds"))
write_csv(df, paste0(data, "df_prepared.csv"))
```

# Session info

```{r echo=TRUE, message=FALSE, warning=FALSE}
sessionInfo()
```
