---
title: "Communicating spatial frames with whole-body silent gesture"
subtitle: "Model fitting"
author: "Analysis by AĆ"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmarkdown::html_document:
    theme: readable
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
    df_print: paged
    code_folding: hide
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '3'
  html_notebook:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is the third script in the spatial-frames pipeline. It fits every model and caches each to its own file in `models/`. Nothing is interpreted here; the fitted objects are the output. Estimates, tests, ROPE decisions, and figures are all in `4_results.Rmd`.

All models are Bayesian multilevel Bernoulli (logit) models fit with `brms`. The family is fixed by design — both outcomes are binary — so there is no likelihood-selection step. Priors, contrasts, and the random-effects structure follow the preregistration.

**The models.**

*Confirmatory.*

-   **H2a** and **H2b** are the strategy hypotheses. Because icon and index are coded as independent, sometimes co-occurring flags rather than one icon-or-index label, they are fit as two separate Bernoulli models — `icon ~ size * distance` and `index ~ size * distance`. One coefficient carries each hypothesis: H2a on the distance slope of the icon model (displaced → more icon), H2b on the distance slope of the index model (proximate → more index).
-   **Model B** (`correct_target ~ size * distance + icon_c + index_c`) carries H1 (the size slope, complete < element) and the no-difference component of H2 (the distance slope, tested against the ROPE). The strategy flags enter centered (±0.5) as covariates.
-   **Model B, no strategy** (`correct_target ~ size * distance`) is the registered robustness fit that recovers the total size effect without conditioning on strategy.
-   **Registered Model A** (`rep_strategy ~ size * distance`) is fit on the subset of productions coded as exactly one of icon or index (`rep_strategy` = 1 for index, 0 for icon). This is the single-binary strategy model as originally preregistered, retained as a robustness check where its icon-or-index assumption holds.

*Exploratory (post hoc, not preregistered).*

-   **Third strategy** (`prop ~ size * distance`): the unregistered third strategy, modeled as its own Bernoulli exactly as icon and index are, to test whether its use depends on the design.
-   **Co-occurrence indicator** (`both ~ size * distance`): a Bernoulli on whether icon and index were used together — the headline test of whether layering depends on condition.
-   **Configuration** (`strategy_combo ~ size * distance`, categorical): the four-way partition `neither / icon_only / index_only / both` with `neither` as reference, giving the full picture of how strategy configuration shifts across the design. Random intercepts only, since a full random-slope structure across three linear predictors is not warranted for an exploratory model.

**Specifications common to the confirmatory models.**

-   `size` and `distance` use unweighted **±0.5 sum coding** (element / proximate = −0.5, complete / displaced = +0.5), so each lower-order term is a main effect at the midpoint of the other factor and the intercept sits at the center of the design.
-   **Priors** (preregistered): Normal(0, 1.5) on fixed slopes, Normal(0, 2) on the intercept, Exponential(1) on random-effect SDs, LKJ(2) on random-effect correlations.
-   **Random effects** are maximal — by-producer and by-guesser intercepts and `size * distance` slopes, plus a by-item intercept — with a preregistered **fallback ladder** applied when a model does not converge (R-hat ≤ 1.01, zero divergent transitions): drop the correlation terms, then the interaction slope, then the main-effect slopes, guesser before producer. `fit_ladder()` fits from the maximal structure down and returns the first level that converges.
-   Four chains, 10,000 iterations, 4,000 warmup, `adapt_delta = 0.95`. 

**Caching.** Each model is saved to `models/<name>.rds` and reloaded on re-knit. Set `refit_all <- TRUE` (or delete `models/`) to refit — the cache keys on the filename, not the data, so switch it on when the real annotation replaces the simulated file.

# Data preparation

## Source setup

```{r source setup, echo = TRUE, message=FALSE, warning = FALSE}

########## folders ##########
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

parentfolder <- dirname(getwd())

data          <- paste0(parentfolder, '/data/')
models        <- paste0(parentfolder, '/models/')
plots         <- paste0(parentfolder, '/plots/')
scripts       <- paste0(parentfolder, '/scripts/')

if (!dir.exists(models)) dir.create(models, recursive = TRUE)

#################### packages ####################
library(tidyverse)
library(knitr)

# Bayesian
library(brms)
library(posterior)     # rhat / draws
library(bayesplot)     # divergence count

# use all available cores for parallel computing
options(mc.cores = parallel::detectCores())

# If cmdstanr is installed it is faster; uncomment to use it globally.
# options(brms.backend = "cmdstanr")

######## refit switch ########
# FALSE = load cached fits from models/ when present (fast re-knit).
# TRUE  = ignore the cache and refit everything. Turn ON when the real
#         annotation replaces the simulated file (cache keys on filename).
refit_all <- FALSE
```

## Load prepared data

```{r load data, echo=TRUE, message=FALSE, warning=FALSE}
df <- readRDS(paste0(data, "df_prepared.rds")) %>%
  mutate(item = story_id)     # 16-level item id for the by-item random effect
```

## Contrasts and derived variables

`size` and `distance` get unweighted ±0.5 sum coding (level 1 = −0.5, level 2 = +0.5). With levels ordered element/complete and proximate/displaced, a positive slope means higher for complete and for displaced, respectively. The strategy flags are centered to ±0.5 for Model B, `both` marks icon-index co-occurrence, and `df_single` is the single-coded subset for the registered Model A.

```{r contrasts, echo=TRUE, message=FALSE, warning=FALSE}
# --- ±0.5 sum coding (equivalent to MASS::contr.sdif(2) for a 2-level factor) ---
contrasts(df$size)     <- matrix(c(-0.5, 0.5), nrow = 2,
                                 dimnames = list(NULL, "compl"))  # element(-) / complete(+)
contrasts(df$distance) <- matrix(c(-0.5, 0.5), nrow = 2,
                                 dimnames = list(NULL, "displ"))  # proximate(-) / displaced(+)

# --- derived variables ---
df <- df %>%
  mutate(
    icon_c  = icon  - 0.5,                       # centered strategy covariates (±0.5)
    index_c = index - 0.5,
    both    = as.integer(icon == 1 & index == 1) # icon-index co-occurrence indicator
  )

# --- single-coded subset for the registered single-binary Model A ---
df_single <- df %>%
  filter(strategy_combo %in% c("icon_only", "index_only")) %>%
  droplevels() %>%
  mutate(rep_strategy = as.integer(strategy_combo == "index_only"))  # 1 = index, 0 = icon

# re-apply contrasts after subsetting (droplevels can drop the attribute)
contrasts(df_single$size)     <- matrix(c(-0.5, 0.5), nrow = 2,
                                        dimnames = list(NULL, "compl"))
contrasts(df_single$distance) <- matrix(c(-0.5, 0.5), nrow = 2,
                                        dimnames = list(NULL, "displ"))
```

# Priors, random-effects ladder, and helpers

```{r helpers, echo=TRUE, message=FALSE, warning=FALSE}
#################### priors (preregistered) ####################
priors_base <- c(
  set_prior("normal(0, 1.5)", class = "b"),
  set_prior("normal(0, 2)",   class = "Intercept"),
  set_prior("exponential(1)", class = "sd")
)
priors_cor <- c(priors_base, set_prior("lkj(2)", class = "cor"))   # + RE correlations

#################### preregistered RE fallback ladder ####################
# maximal -> simplest: drop correlations, then the interaction slope, then the
# main-effect slopes (by-guesser before by-producer). 'cor' flags the level
# that still has correlation parameters (so it gets the LKJ prior).
re_ladder <- list(
  list(re = "(1 + size * distance | producer) + (1 + size * distance | guesser) + (1 | item)", cor = TRUE),
  list(re = "(1 + size * distance || producer) + (1 + size * distance || guesser) + (1 | item)", cor = FALSE),
  list(re = "(1 + size + distance || producer) + (1 + size + distance || guesser) + (1 | item)", cor = FALSE),
  list(re = "(1 + size + distance || producer) + (1 | guesser) + (1 | item)", cor = FALSE),
  list(re = "(1 | producer) + (1 | guesser) + (1 | item)", cor = FALSE)
)

#################### diagnostics ####################
fit_diagnostics <- function(fit) {
  dr   <- posterior::as_draws(fit)
  rh   <- suppressWarnings(max(posterior::summarise_draws(dr, "rhat")$rhat, na.rm = TRUE))
  np   <- bayesplot::nuts_params(fit)
  ndiv <- sum(np$Value[np$Parameter == "divergent__"])
  tibble(
    rhat_max  = round(rh, 4),
    divergent = ndiv,
    converged = is.finite(rh) && rh <= 1.01 && ndiv == 0
  )
}
is_converged <- function(fit) isTRUE(fit_diagnostics(fit)$converged)

#################### fit the RE ladder ####################
# Returns the first RE level that converges; otherwise the simplest, with a
# warning. Fits lazily: stops at the first convergent level.
fit_ladder <- function(outcome, fixed, data,
                       family = bernoulli(),
                       ladder = re_ladder,
                       iter = 10000, warmup = 4000, seed = 2137) {
  chosen <- NULL
  for (i in seq_along(ladder)) {
    f  <- as.formula(paste(outcome, "~", fixed, "+", ladder[[i]]$re))
    pr <- if (ladder[[i]]$cor) priors_cor else priors_base
    message("  RE level ", i, " ...")
    cand <- brm(
      formula = f, data = data, family = family, prior = pr,
      chains = 4, iter = iter, warmup = warmup, seed = seed,
      control = list(adapt_delta = 0.95, max_treedepth = 12),
      refresh = 0
    )
    if (is_converged(cand)) {
      message("  -> converged at RE level ", i)
      return(cand)
    }
    chosen <- cand
  }
  warning(outcome, ": no RE level fully converged; keeping simplest (level ",
          length(ladder), ").")
  chosen
}

#################### cache wrapper: one .rds per model ####################
get_model <- function(tag, fit_fun) {
  path <- paste0(models, tag, ".rds")
  if (!refit_all && file.exists(path)) {
    message("Loading cached model: ", tag)
    return(readRDS(path))
  }
  message("Fitting model: ", tag)
  fit <- fit_fun()
  saveRDS(fit, path)
  fit
}
```

# Model fitting

## Confirmatory models

The two strategy Bernoullis (H2a, H2b), Model B (H1 + H2 no-difference), its no-strategy robustness fit, and the registered single-binary Model A on the single-coded subset. Each runs the RE fallback ladder and is cached to its own file.

```{r fit_confirmatory, echo=TRUE, message=TRUE, warning=FALSE, results='hide'}
# H2a: icon use
m_icon <- get_model("m_icon", function()
  fit_ladder("icon", "size * distance", df))

# H2b: index use
m_index <- get_model("m_index", function()
  fit_ladder("index", "size * distance", df))

# Model B: communicative success (H1 + H2 no-difference), strategy as covariates
m_success <- get_model("m_success", function()
  fit_ladder("correct_target", "size * distance + icon_c + index_c", df))

# Model B robustness: total size effect, no strategy covariates
m_success_nostrat <- get_model("m_success_nostrat", function()
  fit_ladder("correct_target", "size * distance", df))

# Registered single-binary Model A, on the single-coded subset (1 = index)
m_repstrat_registered <- get_model("m_repstrat_registered", function()
  fit_ladder("rep_strategy", "size * distance", df_single))

models_confirmatory <- list(
  icon                = m_icon,
  index               = m_index,
  success             = m_success,
  success_nostrat     = m_success_nostrat,
  repstrat_registered = m_repstrat_registered
)
```

## Exploratory models

Post hoc, not preregistered: the co-occurrence indicator (headline test of condition-dependent layering) and the categorical configuration model (the full four-way partition). The categorical model uses random intercepts only.

```{r fit_exploratory, echo=TRUE, message=TRUE, warning=FALSE, results='hide'}
# third strategy (unregistered): prop use
m_prop <- get_model("m_prop", function()
  fit_ladder("prop", "size * distance", df))

# icon-index co-occurrence indicator
m_both <- get_model("m_both", function()
  fit_ladder("both", "size * distance", df))

# four-way configuration (neither / icon_only / index_only / both), reference = neither
m_combo <- get_model("m_combo", function()
  brm(
    formula = strategy_combo ~ size * distance +
      (1 | producer) + (1 | guesser) + (1 | item),
    data = df, family = categorical(),
    prior = priors_base,
    chains = 4, iter = 10000, warmup = 4000, seed = 2137,
    control = list(adapt_delta = 0.95, max_treedepth = 12),
    refresh = 0
  ))

models_exploratory <- list(prop = m_prop, both = m_both, combo = m_combo)
```

# Convergence summary

R-hat should be ≤ 1.01 and divergent transitions should be 0 for every model. Any `FALSE` in `converged` means the ladder was exhausted (for a confirmatory model) or the categorical model needs a second look before its estimates are used in `4_results.Rmd`.

```{r convergence, echo=TRUE, message=FALSE, warning=FALSE}
all_fits <- c(models_confirmatory, models_exploratory)

convergence_summary <- imap_dfr(all_fits, function(fit, nm)
  bind_cols(model = nm, n_obs = nrow(fit$data), fit_diagnostics(fit)))

convergence_summary
```

```{r convergence_flag, echo=TRUE, message=FALSE, warning=FALSE}
flagged <- convergence_summary$model[!convergence_summary$converged]
if (length(flagged) == 0) {
  message("All models converged (R-hat <= 1.01, 0 divergences).")
} else {
  warning("Look again: ", paste(flagged, collapse = ", "))
}
```

# Session info

```{r echo=TRUE, message=FALSE, warning=FALSE}
sessionInfo()
```
